Operator evolution from the similarity renormalization group and the Magnus expansion

### 2019-09-09 Notes/Operator_evolution/operator_evolution.tex
Updated operator_evolution notes
Added bullets to introduction and further organized things.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% File: operator_evolution.tex
%
% Author: A. J. Tropiano (tropiano.4@osu.edu)
% Date: August 23, 2019
%
% Draft of paper on SRG operator evolution and the Magnus expansion.
%
% Revision history:
%	08/28/19 --- Started outlining and added pieces from old notes.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\noindent{%
\textbf{Background on modern nuclear potentials.}
}
\\
-- Wide range of NN potentials: how do they change under SRG transformations?
\\
-- Implementation to many-body calculations and SRG decoupling.
\\
-- Motivate long list of unaddressed questions: regulator, generator dependence, universality, high cutoffs, the Magnus expansion.
\\
\textbf{SRG formalism}
\\
-- The SRG decouples low- and high-momentum scales by applying a continuous unitary transformation $U(s)$ where $s=0 \rightarrow \infty$ is the flow parameter.
\\
-- The `dressed' or evolved operator is given by
%
\begin{eqnarray}
	\label{eq:srg_operator}
	O(s) = U(s) O(0) U^{\dagger}(s),
\end{eqnarray}
%
where $O(0)$ corresponds to the `bare' operator.
\\
-- Because $U(s)$ is unitary, the observables of the operator are preserved.
\\
-- In practice, the unitary transformation U(s) is not explicitly solved for; the evolved operator is given by a differential flow equation which is obtained by taking the derivative of Eqn. (\ref{eq:srg_operator}),
%
\begin{eqnarray}
	\label{eq:srg_flow}
	\frac{dO(s)}{ds} = [\eta(s), O(s)],
\end{eqnarray}
%
where $\eta(s)=\frac{dU(s)}{ds} U^{\dagger}(s) = -\eta^{\dagger}(s)$ is the anti-hermitian SRG generator.
\\
-- The generator is defined as a commutator,
%
\begin{eqnarray}
	\label{eq:srg_generator}
	\eta(s) = [G, H(s)],
\end{eqnarray}
%
where $G$ specifies the type of flow or form of the decoupled operator.
\\
-- To drive the operator to band-diagonal form, set $G=H_D(s)$, the diagonal of the Hamiltonian.
\\
-- This choice was implemented by Wegner in condensed matter physics \cite{Wegner:1994ab}.
\\
-- For notational convenience, we write the Wegner choice without the $s$ dependence in the rest of the paper.
\\
-- In a similar option used in nuclear physics, $G$ is set to the relative kinetic energy, $T_{rel}$, which also drives to band-diagonal form.
\\
-- In this paper, we consider both choices.
\\
\textcolor{red}{%
-- Add block-diagonal generator.
}
\\
-- Generally the flow equation (\ref{eq:srg_flow}) is solved up to some finite value of $s$ with a high-order ODE solver.
\\
-- It is convenient to define $\lambda \equiv s^{-1/4}$ which roughly measures the width of the band-diagonal in the decoupled operator.
\\
\textbf{Operator evolution}
\\
-- State how a potential and wave function changes: how does this affect other operators?
\\
-- How operators evolve from band- and block-diagonal SRG transformations.
\\
-- Operator evolution for different potentials (regulators, chiral order, etc.)

@@

\noindent{%
-- Comparison of potential evolution with different regulators, orders, generators.
}
\\
-- Universality.
\\
-- Discussion of high cutoffs, block-diagonal generator at high cutoffs, and how it handles spurious bound states.
\\
-- Use high cutoffs to transition to Magnus test problem.
\\
-- Example figure. EM N$^3$LO potential.
%
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_operators/potential_contours_kvnn10_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_operators/potential_contours_kvnn10_3S1_Block-diag2,00}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_operators/potential_contours_kvnn10_3S1_Block-diag3,00}%
	}
	\caption{Matrix elements of the Entem-Machleidt N$^3$LO non-local potential $V_{\lambda}(k, k')$ SRG-evolving in $\lambda$ right to left under transformations with the Wegner generator (a) and block-diagonal generators decoupling at $\Lambda=2$ and $3$ fm$^{-1}$ (b and c).}
	\label{potential_contours_kvnn10_3S1}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% File: operator_evolution.tex
%
% Author: A. J. Tropiano (tropiano.4@osu.edu)
% Date: August 23, 2019
%
% Draft of paper on SRG operator evolution and the Magnus expansion.
%
% Revision history:
%	08/28/19 --- Started outlining and added pieces from old notes.
%	09/09/19 --- Added more bullets to the introduction. Further organization.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\noindent{%
\textbf{Background on modern nuclear potentials.}
}
\\
-- Wide range of NN potentials.
%
\begin{itemize}
	\item Chiral EFT background.
	\item Different potential matrix elements but give same $S$-matrix.
\end{itemize}
%
-- Implementation to many-body calculations and SRG decoupling.
%
\begin{itemize}
	\item Strong coupling between low- and high-momentum matrix elements in realistic potentials.
	\item Very difficult to implement these interactions in many-body methods using basis expansions. Matrix dimension becomes too large for accurate calculations.
	\item RG transformations are used to soften the interaction to make many-body methods feasible. One such method, the SRG, also preserves observables from unitarity.
	\item How do different interactions change under SRG transformations?
\end{itemize}
%
\textbf{SRG formalism}
\\
-- The SRG decouples low- and high-momentum scales by applying a continuous unitary transformation $U(s)$ where $s=0 \rightarrow \infty$ is the flow parameter.
\\
-- The `dressed' or evolved operator is given by
%
\begin{eqnarray}
	\label{eq:srg_operator}
	O(s) = U(s) O(0) U^{\dagger}(s),
\end{eqnarray}
%
where $O(0)$ corresponds to the `bare' operator.
\\
-- Because $U(s)$ is unitary, the observables of the operator are preserved.
\\
-- In practice, the unitary transformation U(s) is not explicitly solved for; the evolved operator is given by a differential flow equation which is obtained by taking the derivative of Eqn. (\ref{eq:srg_operator}),
%
\begin{eqnarray}
	\label{eq:srg_flow}
	\frac{dO(s)}{ds} = [\eta(s), O(s)],
\end{eqnarray}
%
where $\eta(s)=\frac{dU(s)}{ds} U^{\dagger}(s) = -\eta^{\dagger}(s)$ is the anti-hermitian SRG generator.
\\
-- The generator is defined as a commutator, $\eta(s) = [G, H(s)]$, where $G$ specifies the type of flow or form of the decoupled operator.
\\
-- To drive the operator to band-diagonal form, set $G=H_D(s)$, the diagonal of the Hamiltonian.
\\
-- This choice was implemented by Wegner in condensed matter physics \cite{Wegner:1994ab}.
\\
-- In a similar option used in nuclear physics, $G$ is set to the relative kinetic energy, $T_{rel}$, which also drives to band-diagonal form.
\\
-- It is convenient to define $\lambda \equiv s^{-1/4}$ which roughly measures the width of the band-diagonal in the decoupled operator.
\\
-- For block-diagonal decoupling, denoted $G=H_{BD}(s)$, the operator is split into low- and high-momentum sub-blocks by specifying a separation in momentum $\Lambda_{BD}$.
\\
-- These transformations are similar to $V_{low k}$ Lee-Suzuki transformations but keep the high-momentum matrix elements non-zero, although entirely decoupled from the low-momentum sub-block.
\\
-- Generally the flow equation (\ref{eq:srg_flow}) is solved up to some finite value of $s$ with a high-order ODE solver.
\\
-- For notational convenience, we write the generators without the $s$ dependence in the rest of the paper.
\\
\textbf{Universality}
\\
-- The explicit long-range physics should be the same. Decoupling low- and high-energy gives matching low-energy matrix elements. In the NN potential, this means softened NN interactions should have the same low-momentum matrix elements after sufficient decoupling.
\\
-- Add takeaways from Dainton: phase shift equivalence implies matrix element equivalence for $\lambda$ approaching the momenta of phase equivalence. Correct binding energy is critical or the lowest matrix elements will not match.
\\
-- Motivate long list of unaddressed questions: regulator, generator dependence, high cutoffs, the Magnus expansion.
%
\begin{itemize}
	\item Regulator. Functional dependence of regulator. Reasons for implementing each.
	\item Generators. Band- and block-diagonal transformations.
	\item High cutoffs. Cite Nogga, Wendt, and Tews papers.
	\item The Magnus expansion. High cutoffs and connection to IMSRG intruder state problem.
\end{itemize}
%
\textbf{Operator evolution}
\\
-- State how a potential and wave function changes: how does this affect other operators?
\\
-- Operator evolution for different potentials (regulators, chiral order, etc.)
\\
-- How operators evolve from band- and block-diagonal SRG transformations.
\\
\textbf{Overview of sections.}

@@

\noindent{%
-- Comparison of potential evolution with different regulators, orders, generators.
}
\\
-- Universality.
\\
-- Discussion of high cutoffs, block-diagonal generator at high cutoffs, and how it handles spurious bound states.
\\
-- Use high cutoffs to transition to Magnus test problem.
\\
-- Example figure. EM N$^3$LO potential.
%
\begin{figure}[H]
	\centering
	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn10_3S1_Wegner}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn10_3S1_Block-diag2,00}%
	}

	\subfloat[]{%
	\includegraphics[clip,width=0.9\columnwidth]{SRG_potentials/potential_contours_kvnn10_3S1_Block-diag3,00}%
	}
	\caption{Matrix elements of the Entem-Machleidt N$^3$LO non-local potential $V_{\lambda}(k, k')$ SRG-evolving in $\lambda$ right to left under transformations with the Wegner generator (a) and block-diagonal generators decoupling at $\Lambda=2$ and $3$ fm$^{-1}$ (b and c).}
	\label{potential_contours_kvnn10_3S1}
\end{figure}
